Title : The Italian Crisis and Producer Households Debt: a Source of Stability? A Reproducible Research
==============================================================
Stefano Olgiati, Gilberto Bronzini, Alessandro Danovi 
-----------------------------------------------------------------

_University of Bergamo_
-----------------------------------------------------------------

### Legenda and Reproducible Data Source

LEGENDA BANCA D'ITALIA TDB30486_DOMI_ENG.TXT  

> ENTE_SEGN  REPORTING INSTITUTION  
3691030	BANKS, FINANCIAL COMPANIES AND OTHER INSTITUTIONS REPORTING TO THE CCR  

> VOCESOTVOC | PHENOMENA OBSERVED  
351122133	LOAN FACILITIES (EXCLUDING ADJUSTED BAD DEBTS) : CREDIT USED  
351122141	LOAN FACILITIES (EXCLUDING ADJUSTED BAD DEBTS) : NUMBER OF BORROWERS  
035120163	QUARTERLY DEFALUT RATES FOR LOAN FACILITIES : NUMER OF BORROWERS)  
035120363	QUARTERLY DEFAULT RATES FOR LOAN FACILITIES : CREDIT USED  
351121433	QUARTERLY FLOW - ADJUSTED BAD DEBTS AT T FROM PERFORMING LOANS AT T-1: AMOUNT  
351121441	QUARTERLY FLOW - ADJUSTED BAD DEBTS AT T FROM PERFORMING LOANS AT T-1:BORROWERS  

> CLASSE_UTI	TOTAL CREDIT USED (SIZE CLASSES)  
1006	500,000 EUROS AND MORE  
1005	FROM >= 125,000 TO < 500,000 EUROS  
1004	LESS THAN 125,000 EUROS  
9904	TOTAL (>=0)  

> SETCON	CUSTOMER SECTORS AND SEGMENTS OF ECONOMIC ACTIVITY  
7015	CONSUMER HOUSEHOLDS  
0026	FINANCIAL COMPANIES OTHER THAN MONETARY FINANCIAL INSTITUTIONS  
0001	GENERAL GOVERNMENT  
7006	NON-FINANCIAL CORPORATIONS  
7009	NON-PROFIT INSTITUTIONS SERVING HOUSEHOLDS AND UNCLASSIFIABLE UNITS  
7010	PRODUCER HOUSEHOLDS  
4015	TOTAL RESIDENT SECTOR EXCLUDING MFIS  

Download data from Banca d'Italia Public Online Database BIP Online: URI "http://bip.bancaditalia.it/4972unix/pls/4972unix_cartridge/scaricapkg.frameS?lang=ENG&p_tavola=TDB30486&da_dove=&dat_agg=&P1_VARPIANO_VAL=035120163]00001]3691030]&P1_VARCELLA=VALORE]&P1_VARRIGA=DATA_OSSER]&P1_VARCOLONNA=CLASSE_UTI]&P1_VARPIANO=VOCESOTVOC]SETCON]ENTE_SEGN]&P1_IDENTREPORT=2248&P1_OPZIONI=00D&pNumCol=4&pNumRow=25&pPagUtente=NO&pvFuoriPiano=&id_sessione=2905&pvUtenza=&pvValuta="

Choose option:
2. Data: whole table (*)

Installing R packages for analysis and synthesis

```{r}
require(data.table)
require(ggplot2)
require(MASS)
require(markdown)
require(knitr)
library(Quandl)
```

```{r hide}
input.data = read.csv("LOCAL REPO DIRECTORY AFTER CLONING REPO FROM GITHUB")
head(input.data)
tail(input.data)
```


### Parsing data, creating the data frame and calculating ratios for Households and Non-financial Corporations Credit and Bad Debts in Italy: 1996-2013

```{r}
raw.data = data.frame("SEGMENT"=input.data$SETCON,
                      "PHENOMENA" = input.data$VOCESOTVOC, 
                      "SIZE_CLASS" = input.data$CLASSE_UTI, 
                      "DATE" = as.Date(input.data$DATA_OSSER, "%d/%m/%Y"),
                      "VALUE" = input.data$VALORE)
total.data = raw.data[raw.data$SIZE_CLASS %in% c("9904"),]
```

Extracting data on producer households

```{r}
households.producer.data = total.data[total.data$SEGMENT %in% c("7010"),]
loan.amount.producer.data= households.producer.data[households.producer.data$PHENOMENA %in% c("351122133"),]
loan.borrowers.producer.data= households.producer.data[households.producer.data$PHENOMENA %in% c("351122141"),]
default.amount.producer.data= households.producer.data[households.producer.data$PHENOMENA %in% c("351121433"),]
default.borrowers.producer.data= households.producer.data[households.producer.data$PHENOMENA %in% c("351121441"),]
```

Extracting data on non-financial corporations (NFC)

```{r}
NFC.data = total.data[total.data$SEGMENT %in% c("7006"),]
loan.amount.NFC.data= NFC.data[NFC.data$PHENOMENA %in% c("351122133"),]
loan.borrowers.NFC.data= NFC.data[NFC.data$PHENOMENA %in% c("351122141"),]
default.amount.NFC.data= NFC.data[NFC.data$PHENOMENA %in% c("351121433"),]
default.borrowers.NFC.data= NFC.data[NFC.data$PHENOMENA %in% c("351121441"),]
```

Making a new data frame with producer households and NFC

```{r}
data.df.reverse = data.frame("Date"=loan.amount.producer.data$DATE, 
                             "Credit_Amount_Producer"=loan.amount.producer.data$VALUE, 
                             "Credit_Borrowers_Producer"=loan.borrowers.producer.data$VALUE, 
                             "Default_Amount_Producer"=default.amount.producer.data$VALUE,
                             "Default_Borrowers_Producer"=default.borrowers.producer.data$VALUE,
                              "Credit_Amount_NFC"=loan.amount.NFC.data$VALUE, 
                             "Credit_Borrowers_NFC"=loan.borrowers.NFC.data$VALUE, 
                             "Default_Amount_NFC"=default.amount.NFC.data$VALUE,
                             "Default_Borrowers_NFC"=default.borrowers.NFC.data$VALUE)
```

Defining the ratio of per capita credit and bad debts between NFC and producer households

$credit.NFC.percapita=Credit Amount NFC/Credit Borrowers NFC$
$default.NFC.percapita=Default Amount NFC/Default Borrowers NFC$
$credit.producers.percapita=Credit Amount Producer/Credit Borrowers Producer$
$default.producers.percapita=Default Amount Producer/Default Borrowers Producer$

Defining the ratio of per capita credit and bad debts between NFC and producer households

$$credit.ratio=credit.NFC.percapita/credit.producers.percapita.$$
$$baddebt.ratio=baddebt.NFC.percapita/baddebt.producers.percapita$$

Calculating the ratio of per capita credit between NFC and producer households

```{r}
credit.producers.percapita = (data.df.reverse[,"Credit_Amount_Producer"] / data.df.reverse[,"Credit_Borrowers_Producer"])
credit.NFC.percapita = (data.df.reverse[,"Credit_Amount_NFC"] / data.df.reverse[,"Credit_Borrowers_NFC"])
credit.ratio = credit.NFC.percapita / credit.producers.percapita
```

Calculating the ratio of per capita bad debts between NFC and producer households

```{r}
baddebt.producers.percapita = (data.df.reverse[,"Default_Amount_Producer"] / data.df.reverse[,"Default_Borrowers_Producer"])
baddebt.NFC.percapita = (data.df.reverse[,"Default_Amount_NFC"] / data.df.reverse[,"Default_Borrowers_NFC"])
baddebt.ratio = baddebt.NFC.percapita / baddebt.producers.percapita
```

Calculating the average default rates and standard deviation

```{r}
defaultrate.producers = (data.df.reverse[,"Default_Amount_Producer"] / data.df.reverse[,"Credit_Amount_Producer"])
defaultrate.NFC = (data.df.reverse[,"Default_Amount_NFC"] / data.df.reverse[,"Credit_Amount_NFC"])
summary(defaultrate.producers)
variationCoefficient.producers=sd(defaultrate.producers) / mean(defaultrate.producers)
variationCoefficient.producers
summary(defaultrate.NFC)
variationCoefficient.NFC=sd(defaultrate.NFC) / mean(defaultrate.NFC)
variationCoefficient.NFC

qqplot(defaultrate.producers, defaultrate.NFC)
abline(c(0,0), c(1,1), col='red')
```



Making syhthesis data frame of per capita credit, per capita bad debt and ratios

```{r}
syn.data.reverse.df = data.frame("Date"=loan.amount.producer.data$DATE,
                                       "Per_Capita_Credit_NFC" = credit.NFC.percapita,
                                       "Per_Capita_Credit_Producers" = credit.producers.percapita,
                                       "Credit_Ratio"= credit.ratio,
                                       "Per_Capita_Baddebts_NFC" = baddebt.NFC.percapita,
                                       "Per_Capita_Baddebts_Producers" = baddebt.producers.percapita,
                                       "Baddebt_Ratio"= baddebt.ratio)
```

Reversing data frame: head = begin_date and tail = end_date

```{r}
syn.data.df = syn.data.reverse.df[order(loan.amount.producer.data$DATE),]
syn.data.dtable = as.data.table(syn.data.df)
```

### Summary of parsed Data

```{r}
summary(syn.data.dtable)
```

Plots
---------------------------------------

Plotting the time series of credit.ratio with quadratic fit

```{r}
lm.fit.credit.ratio = lm(credit.ratio ~ poly(loan.amount.producer.data$DATE, 5))
qplot(loan.amount.producer.data$DATE, credit.ratio,
      geom = c("point", "smooth"),
      method = "rlm",
      formula = y ~ poly(x, 5),
      main="Per capita Credit Ratio in Italy: 1996-2013 \n Non-financial Corporatio/Producer Households",
      xlab = "Year",
      ylab = "Per capita Credit Ratio")
confint(lm.fit.credit.ratio)
```

Plotting the time series of baddebt.ratio with quadratic fit

```{r}
lm.fit.baddebt.ratio = lm(baddebt.ratio ~ poly(loan.amount.producer.data$DATE, 5))
qplot(loan.amount.producer.data$DATE, baddebt.ratio,
      geom = c("point", "smooth"),
      method = "rlm",
      formula = y ~ poly(x, 5),
      main="Per capita Bad Debts Ratio in Italy: 1996-2013 \n Non-financial Corporatio/Producer Households",
      xlab = "Year",
      ylab = "Per capita Bad Debt Ratio")
confint(lm.fit.baddebt.ratio)
```

QQ plot of credit.ratio, baddebt.ratio

```{r}
qqplot(credit.ratio, 
       baddebt.ratio, 
       cex.main=0.85, 
       main="QQ Plot \n Per capita Credit/Bad Debt  Ratio in Italy: 1996-2013 \n Non-financial Corporations/Producer Households")
abline(c(0,0), c(1,1), col='red')
```

Summary Statistics, Robustness and Confidence Intervals
------------------------------------------------

```{r}
summary(lm.fit.credit.ratio)
summary(lm.fit.baddebt.ratio)
```


*Copyright Stefano Olgiati  2013-12-08*